### 2020-06-16 Format.Rmd
Partitioning of data

---
title: "Federated Imputation"
author: 
- Thom Volker
- Utrecht University
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SET UP

## TO DO

- [x] Dataset: Boys data in mice
- [x] Partition data in 5 subsets
- [ ] Create synthetic versions of the five subsets
- [ ] Merge 5 synthetic subsets into one central synthetic dataset
- [ ] Impute the central synthetic dataset
- [ ] Deconstruct the m imputed sets into the 5 synthetic sets
- [ ] Match the corresponding synthetic imputation to the observed imputation in all of the (m by 5) synthetic imputed sets

## Evaluation plan - when are we satisfied?
- [ ] Has algorithm converged (obv synthetic version --> evt later obv de real data version)
- [ ] Hanne's suggestion for tracking a model/data parameter based on the real data
- [ ] Als bias, CI width en Coverage of the 95% CI about the parameter (e.g. beta or mean) ok is.

## Workflow

Proof of Concept: Incomplete boys data as starting point and assess:

1. convergence?
2. whether the regression model based on federated imputations differs from the regression model based on the unfederated data. 
3. 1000 sims.

IF PoC successful --> jeej

1. Full blown sim with complete(mice(boys, m = 1, seed = 123)) as TRUE/population input
2. Bias? CI width? Coverage?

# REMARKS

1. Randomly subsetting the data leads to "randomly equivalent" datasets. However, when pooling results over multiple nodes, it might be that there are qualitative differences between nodes, which might complicate the imputation process.
2. In practice, some variables may be entirely missing, which differs from the current situation in which all datasets consist of the exact same variables.
3. Start with a single complete datasets, cut it in five, create a synthetic version of all and run the analyses on this synthetic dataset.
4. If I recall it correctly, Stef said that we simply synthesize the individual (partitioned) datasets, and that we could simply Rbind these. However, this would skip the point of basing the imputations on the collection of the 5 datasets together. If this was the case, we could simply use the data as if it is in the nodes, impute the individual sets with reference to the data in the other nodes, and analyse the imputed datasets, and pool the results.
5. Synthpop treats missingness as a separate category. For categorical variables, this is fairly straightforward: the category NA is added to the existing categories. In case of continuous data, the first step involves the "prediction" of whether or not a variable is missing, if it is predicted to be missing, the value will be NA, if the observation is not predicted to be missing, a synthetic value will be generated.

# Literature to read
Abowd, J. M., & Lane, J. (2004). New approaches to confidentiality protection: Synthetic data, remote access and research data centers. In International workshop on privacy in statistical databases (pp. 282-289). Springer, Berlin, Heidelberg.

- Why and how synthetic data can be used so that no information can be traced back to any individual.

Nowok, B. synthpop: An R package for generatingsynthetic versions of sensitive microdatafor statistical disclosure control. Retrieved from [link](https://www.unece.org/fileadmin/DAM/stats/documents/ece/ces/ge.46/20150/Paper_24_bnowok_synthpop.pdf).

Patki, N., Wedge, R., & Veeramachaneni, K. (2016, October). The synthetic data vault. In 2016 IEEE International Conference on Data Science and Advanced Analytics (DSAA) (pp. 399-410). IEEE. Retrieved from [link](https://www.computer.org/csdl/pds/api/csdl/proceedings/download-article/12OmNwx3Q7S/pdf?token=eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpc3MiOiJjc2RsX2FwaSIsImF1ZCI6ImNzZGxfYXBpX2Rvd25sb2FkX3Rva2VuIiwic3ViIjoiYW5vbnltb3VzQGNvbXB1dGVyLm9yZyIsImVtYWlsIjoiYW5vbnltb3VzQGNvbXB1dGVyLm9yZyIsImV4cCI6MTU5MjIyMDg2Nn0.JhJGad-Fwo1N1vYkh737g_UwZ_jLNXSn1lcTCyVhaJU).

- Example of how to generate synthetic values, might give an idea of how this process works.

Reiter, J. P. (2005). Releasing multiply imputed, synthetic public use microdata: An illustration and empirical study. Journal of the Royal Statistical Society: Series A (Statistics in Society), 168(1), 185-205.

[Synthpop Website.](https://synthpop.org.uk/index.html)