### 2020-12-04 mice_synthesizing/mice_updates.Rmd

---
title: "Mice udpates for synthetic data"
author: "Thom Volker"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
bibliography: federated_imp.bib
csl: "apa-6th-edition.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions that must be added for working with synthetic data

1. Correct pooling function
2. Correct plot function (that uses the observed data and the imputed data)
3. Passive imputation apparently uses the observed data if available, which is not what we want when synthesizing data. 
4. Function to check the confidentiality of the actual data
5. Update parlmice (can only impute missing data, and does not work with the where matrix).