### 2021-04-27 mice_synthesizing/mice_append.Rmd

---
title: "Mice - append empty data.frame"
author: "Thom Volker"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

# Load required packages

```{r}
library(mice) # imputations
library(tidyverse) # tidy data
library(modelr) # bootstrap
library(magrittr) # pipe
library(furrr) # parallel mapping
```

# Set seed

```{r}
set.seed(123) # seed for reproducibility
```

# Create a single completely observed dataset

```{r}
comp_meth <- make.method(boys)
comp_meth['bmi'] <- "~I(wgt/(hgt/100)^2)"

comp_pred <- make.predictorMatrix(boys)
comp_pred[c("wgt", "hgt"), "bmi"] <- 0

truth <- boys %>% 
  mice(seed = 123, 
       comp_meth, 
       m = 1, 
       print = FALSE) %>% 
  complete()
```

# Create the true model

```{r}
model <- function(data) lm(wgt ~ age + hgt, data)

truemodel <- truth %>% model

coefs <- coef(truemodel)

```

# Initialize parallel processing

```{r}
plan(multisession)
```

# Specify number of simulations - 100 for now

```{r}
nsim <- 100
```

# CART method and predictor matrix for synthesizing

```{r}
# cart method, all variables are imputed by means of cart
cart <- rep("cart", ncol(truth))
names(cart) <- colnames(truth)
cart['bmi'] <- "~I(wgt / (hgt/100)^2)"

# alter the predictor matrix such that imputations for bmi do not flow
# back into the predictions for wgt and hgt
pred <- make.predictorMatrix(truth)
pred[c("wgt", "hgt"), "bmi"] <- 0
```

# Bootstrap set of the boys data

```{r}
bootstrap_boys <- bootstrap(truth, nsim) %$% strap %>% map(as.data.frame)
```

# Simulate

```{r, results = F, cache = TRUE, cache.lazy = FALSE}
# 1 iteration, m = 5
boot_cart_append_maxit1_cp_min3 <- bootstrap_boys %>%
  future_map(function(x) {
    x %>% 
      bind_rows(data.frame(matrix(NA, nrow(.), ncol(.), dimnames = list(NULL, colnames(.))))) %>%
      mice(m = 5,
           maxit = 1,
           method = cart,
           minbucket = 3, 
           cp = 1e-08,
           predictorMatrix = pred,
           print = F)
  }, .options = future_options(seed = as.integer(123)), .progress = FALSE)

# 50 iterations, m = 5
boot_cart_append_maxit50_cp_min3 <- bootstrap_boys %>%
  future_map(function(x) {
    x %>% 
      bind_rows(data.frame(matrix(NA, nrow(.), ncol(.), dimnames = list(NULL, colnames(.))))) %>%
      mice(m = 5,
           maxit = 50,
           method = cart,
           minbucket = 3, 
           cp = 1e-08,
           predictorMatrix = pred,
           print = F)
  }, .options = future_options(seed = as.integer(123)), .progress = FALSE)
```

# Pooling rules and coverage function

```{r}
# FULLY SYNTHETIC POOL FUNCTION
pool.syn <- function(mira) {

  if(class(mira)[1] == "mira") { # if the input object is of class mira
    fitlist <- mira %$% analyses # extract the analyses from the mira object
  }
  else {                         # and otherwise, just take the input list
    fitlist <- mira
    }

  m <- length(fitlist)           # number of imputations

  pooled <- fitlist %>% 
    map_dfr(broom::tidy) %>%     # tidy estimates
    group_by(term) %>%           # group per variable
    summarise(est     = mean(estimate),
              bm      = sum((estimate - est)^2) / (m - 1),
              ubar    = mean(std.error^2),
              var_u   = (1 + 1/m) * bm - ubar,
              var     = if_else(var_u > 0, var_u, ubar), # restrict variance to be positive
              df      = max(1, (m - 1) * (1 - ubar / (bm + bm/m))^2), # restrict df > 1
              lower   = est - qt(.975, df) * sqrt(var),
              upper   = est + qt(.975, df) * sqrt(var), .groups = 'drop')
  pooled
}

ci_cov <- function(pooled, true_fit = NULL, coefs = NULL, vars = NULL) {

  if (!is.null(true_fit)) {
    coefs <- coef(true_fit)
    vars   <- diag(vcov(true_fit))
  }

  nsim <- nrow(pooled) / length(coefs)

  pooled %>% mutate(true_coef = rep(coefs, nsim),
                    true_var  = rep(vars, nsim),
                    cover     = lower < true_coef & true_coef < upper) %>%
    group_by(term) %>%
    summarise("True Est" = unique(true_coef),
              "Syn Est"  = mean(est),
              "Bias"     = mean(est - true_coef),
              "True SE"  = unique(sqrt(true_var)),
              "Syn SE"   = mean(sqrt(var)),
              "df"       = mean(df),
              "Lower"    = mean(lower),
              "Upper"    = mean(upper),
              "CIW"      = mean(upper - lower),
              "Coverage" = mean(cover), .groups = "drop")
}
```

# Results of simulation

```{r}
library(kableExtra)

boot_cart_append_maxit1_cp_min3 %>%
  map_dfr(function(x) {
    x %>%
      complete(action = "all") %>%
      map(~model(.x[749:1496,])) %>%
      pool.syn
  }) %>%
  ci_cov(., truemodel)   %>%
  kable(digits = 3,
        caption = "Synthesizing by appending an empty dataframe under the bootstrap samples with 1 iteration") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

boot_cart_append_maxit50_cp_min3 %>%
  map_dfr(function(x) {
    x %>%
      complete(action = "all") %>%
      map(~model(.x[749:1496,])) %>%
      pool.syn
  }) %>%
  ci_cov(., truemodel) %>%
  kable(digits = 3,
        caption = "Synthesizing by appending an empty dataframe under the bootstrap samples with 50 iteration") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

# What happens when running mice for 500 iterations

```{r, cache = TRUE, cache.lazy = FALSE}
iter500 <- boys %>%
  bind_rows(data.frame(matrix(NA, nrow(.), ncol(.), dimnames = list(NULL, colnames(.))))) %>%
  mice(m = 5,
       maxit = 500,
       method = cart,
       minbucket = 3, 
       cp = 1e-08,
       predictorMatrix = pred,
       print = F)

plot(iter500)
```