### 2021-05-13 ads_thesis/simulation_replication.Rmd
simulation replication ADS thesis

---
title: "ADS Thesis simulation"
author: 
- Thom Volker
- Utrecht University
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, results = F, message = F}
library(mice)
library(tidyverse)
library(furrr)
library(magrittr)
library(knitr)
```

# Read in the data

```{r}
data_url <- "https://raw.githubusercontent.com/gerkovink/Synthetic_ADS/main/Data/diabetes.csv"
data <- read_csv(url(data_url), 
                 col_types = cols(Pregnancies = col_integer(), 
                                  Glucose = col_integer(), 
                                  BloodPressure = col_integer(), 
                                  SkinThickness = col_integer(), 
                                  Insulin = col_integer(), 
                                  Age = col_integer(), 
                                  Outcome = col_factor(levels = c("1", "0"))))

data <- data[!data$BMI == 0,]
```

# True model

```{r}
fit <- glm(Outcome ~ Pregnancies + Glucose + BMI, 
           family = binomial(link = "logit"), 
           data = data)

coef(fit)
sqrt(diag(vcov(fit)))
broom::tidy(fit)
```

# Synthetic data model

```{r}
mod1 <- c(Pregnancies              = "pmm", 
          Glucose                  = "norm", 
          BloodPressure            = "norm", 
          SkinThickness            = "pmm", 
          Insulin                  = "pmm", 
          BMI                      = "norm", 
          DiabetesPedigreeFunction = "pmm", 
          Age                      = "pmm", 
          Outcome                  = "logreg.boot")

mod2 <- c(Pregnancies              = "cart", 
          Glucose                  = "cart", 
          BloodPressure            = "cart", 
          SkinThickness            = "cart", 
          Insulin                  = "cart", 
          BMI                      = "cart", 
          DiabetesPedigreeFunction = "cart", 
          Age                      = "cart", 
          Outcome                  = "cart")

pred <- make.predictorMatrix(data)
```

# Simulate

```{r, cache = TRUE, cache.lazy = FALSE}
nsim <- 100
plan(multisession)

syn1 <- future_map(1:nsim, function(x) {
  data %>%
    mice(m = 5,
         method = mod1,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(.), ncol(.)),
         print = FALSE)
}, .options = future_options(seed = as.integer(123)), .progress = FALSE)

syn1_maxit1 <- future_map(1:nsim, function(x) {
  data %>%
    mice(m = 5, 
         maxit = 1,
         method = mod1,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(.), ncol(.)),
         print = FALSE)
}, .options = future_options(seed = as.integer(123)), .progress = FALSE)
```

```{r, cache = TRUE, cache.lazy = FALSE}
syn2 <- future_map(1:nsim, function(x) {
  data %>%
    mice(m = 5,
         method = mod2,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(.), ncol(.)),
         print = FALSE)
}, .options = future_options(seed = as.integer(123)), .progress = FALSE)

syn2_maxit1 <- future_map(1:nsim, function(x) {
  data %>%
    mice(m = 5, 
         maxit = 1,
         method = mod2,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(.), ncol(.)),
         print = FALSE)
}, .options = future_options(seed = as.integer(123)), .progress = FALSE)
```

```{r, cache = TRUE, cache.lazy = FALSE}
syn2_cp <- future_map(1:nsim, function(x) {
  data %>%
    mice(m = 5,
         method = mod2,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(.), ncol(.)),
         cp = 1e-32,
         minbucket = 3,
         print = FALSE)
}, .options = future_options(seed = as.integer(123)), .progress = FALSE)

syn2_cp_maxit1 <- future_map(1:nsim, function(x) {
  data %>%
    mice(m = 5, 
         maxit = 1,
         method = mod2,
         predictorMatrix = pred,
         where = matrix(TRUE, nrow(.), ncol(.)),
         cp = 1e-32,
         minbucket = 3,
         print = FALSE)
}, .options = future_options(seed = as.integer(123)), .progress = FALSE)
```

# Source functions file for pooling

```{r}
source("~/Documents/Federated_imputation/mice_synthesizing/simulations/functions.R")
```

# Synthetic data estimates

```{r, include = F}
library(kableExtra)
```

```{r}
syn1 %>%
  map(function(x) {
    x %$% 
      glm(Outcome ~ Pregnancies + Glucose + BMI, 
          family = binomial(link = "logit"))
    }) %>%
  map_dfr(pool3.syn) %>%
  ci_cov(., fit) %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

syn1_maxit1 %>%
  map(function(x) {
    x %$% 
      glm(Outcome ~ Pregnancies + Glucose + BMI, 
          family = binomial(link = "logit"))
    }) %>%
  map_dfr(pool3.syn) %>%
  ci_cov(., fit) %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r}
syn2 %>%
  map(function(x) {
    x %$% 
      glm(Outcome ~ Pregnancies + Glucose + BMI, 
          family = binomial(link = "logit"))
    }) %>%
  map_dfr(pool3.syn) %>%
  ci_cov(., fit) %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

syn2_maxit1 %>%
  map(function(x) {
    x %$% 
      glm(Outcome ~ Pregnancies + Glucose + BMI, 
          family = binomial(link = "logit"))
    }) %>%
  map_dfr(pool3.syn) %>%
  ci_cov(., fit) %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

```{r}
syn2_cp %>%
  map(function(x) {
    x %$% 
      glm(Outcome ~ Pregnancies + Glucose + BMI, 
          family = binomial(link = "logit"))
    }) %>%
  map_dfr(pool3.syn) %>%
  ci_cov(., fit) %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

syn2_cp_maxit1 %>%
  map(function(x) {
    x %$% 
      glm(Outcome ~ Pregnancies + Glucose + BMI, 
          family = binomial(link = "logit"))
    }) %>%
  map_dfr(pool3.syn) %>%
  ci_cov(., fit) %>%
  kable(digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```