### 2021-09-15 mice_synthesizing/distributional_sim_study.Rmd

---
title: "Untitled"
author: 
- Thom Volker
- Utrecht University
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    highlight: pygments
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulation study -- creating synthetic data with `mice`

# Load packages

```{r}
library(mice)
library(tidyverse)
library(furrr)
library(magrittr)
library(knitr)
```

# Read in the data

```{r}
data_url <- "https://raw.githubusercontent.com/gerkovink/Synthetic_ADS/main/Data/diabetes.csv"
data <- read_csv(url(data_url), 
                 col_types = cols(Pregnancies = col_integer(), 
                                  Glucose = col_integer(), 
                                  BloodPressure = col_integer(), 
                                  SkinThickness = col_integer(), 
                                  Insulin = col_integer(), 
                                  Age = col_integer(), 
                                  Outcome = col_factor(levels = c("0", "1"))))
```

# True model

```{r}
fit <- glm(Outcome ~ BMI + Glucose + Pregnancies, 
           family = binomial(link = "logit"), 
           data = data)

coef(fit)
sqrt(diag(vcov(fit)))
broom::tidy(fit)
```