### 2021-10-27 manuscript/Blind/Submitted/cover letter.tex
Gerko

 \documentstyle{letter}
 \begin{document}
 \pagestyle{empty}

Dear Editor,\\
\newline

We would like to submit the attached manuscript, ”Anonymiced shareable data: Using mice to create and analyze multiply imputed synthetic data sets” for consideration for possible publication in the special issue on \emph{Computational Aspects, Statistical Algorithms and Software in Psychometrics} in \texttt{Psych}. The manuscript outlines a procedure to generate synthetic data sets with \texttt{mice} in {R}. The proposed workflow yields valid univariate and multivariate estimates and allows for valid inference on synthetic data sets generated with \texttt{mice}. We believe that the procedure in our manuscript solves an important problem for applied research and the readership of \texttt{Psych}.

The paper has not been published or accepted for publication elsewhere. It is not under consideration at another journal or at Sociological Methods and Research.

Sincerely, 
\newline\\  
Thom Volker\\
Gerko Vink\\
\newline

Contact Information:\\
Thom Volker\\
Utrecht University\\
Department of Methodology and Statistics\\
PO Box 80140\\
3508 TC Utrecht\\
The Netherlands\\
e-mail: t.b.volker@uu.nl\\

 \end{document}